---
title: "(0) Getting started"
output: html_document
runtime: shiny
---

To create a Shiny app from scratch, navigate to **File** -> **New File**, and select **Shiny Web App...**

All Shiny apps have three components:

1) The **user interface (UI)** sets the layout/appearance of the app.
2) The **server function** contains instructions needed to build the app.
3) Finally, a call to the `shinyApp()` function creates the app from the input server/UI pair.

### Organizing your code

A Shiny app can be written in two ways. By default, RStudio will generate a single `app.R` file that contains all of the components mentioned above. In this case, all required libraries must be loaded and scripts must be sourced inside of the `server()` function.

```{r eval = FALSE}
# app.R
library(shiny)

ui <- fluidPage(plotOutput(outputId = "histogram"))

server <- function(input, output)
    output$histogram <- renderPlot(hist(cars$speed))

shinyApp(ui = ui, server = server)
```

Alternatively, the app can be organized into separate `ui.R` and `server.R` files. An additional `global.R` file can then be used to load libraries and additional scripts. 

```{r eval = FALSE}
# dir/ui.R
fluidPage(plotOutput(outputId = "histogram"))

# dir/server.R
shinyServer(function(input, output)
    output$histogram <- renderPlot(hist(cars$speed))

# dir/global.R
library(shiny)
    
# click on 'Run App' in RStudio run the app via
runApp("dir")
```

Both of the above examples will render the same result. However, for larger projects, it is recommended to resolve to the second option. 

