---
title: "Movies app"
subtitle: "SIB - Introduction to R Shiny"
author: "Phil Cheng, Helena L. Crowell"
date: "May 6th, 2019"
output: 
  learnr::tutorial:
    progressive: true
    theme: flatly
runtime: shiny_prerendered
---

```{r context = "setup", echo = FALSE, message = FALSE, warning = FALSE}
library(dplyr)
library(ggplot2)
library(learnr)
learnr::initialize_tutorial()
knitr::opts_chunk$set(echo = FALSE)
```

```{r context = "data"}
movies <- read.csv(file.path("www", "movies.csv"))
```

## Heads-up

- each of you received one <span style = "color:red"> red </span> and one <span style = "color:green"> green </span> sticky note
- stick these on the back of your laptop:
  - <span style = "color:green"> green </span> when you're **done**
  - <span style = "color:red"> red </span> when you have **questions** or are **stuck**

### &nbsp;

### &nbsp;

- we'll start with **very simple** tasks  
  > use your <span style = "color:green"> green </span> sticky  
  - so you don't get bored
  - so we don't have to awkwardly ask *"are you done?"* ;^)
  
- the tasks will get **very challenging**  
  > use your <span style = "color:red"> red </span> sticky  
    - so you don't get bored
    - so Phil & I can help you out


## Data description

- **movies.csv** contains data from a forum,  
where users collect on-screen body counts for a selection of films,  
and the characters and actors who appear in these films.
- To be counted a kill and/or dead body has to be visible on the screen,  
implied deaths like those in the explosion of the Death Star are not counted.
- The dataset currently contains counts for 545 films from 1949 through 2013.

## Task 0: Creating an app

<div class = "row">
<div class = "col-md-10">
<div style="color:red;font-weight:bold">
1. Create a Shiny app from scratch by navigating to  
**File** > **New File** > **Shiny Web App...**  
(place the file inside the `SIB-Shiny-intro` folder)
2. Launch the app by opening **app.R** and clicking **Run App**
3. Close the app by clicking the stop icon
4. Please remove everything following `ui <-` and `server <-`
</div>
</div>
<div class = "col-md-2">
<div style="color:black; float:right"> `r icon::fa("clock")` 2:00 </div>
</div>
</div>

## Task 1: Loading the data

<div class = "row">
<div class = "col-md-10">
<div style="color:red;font-weight:bold">
1. Set up a `SidebarLayout` with an empty `sidebarPanel`,  
and a `tableOutput` in its `mainPanel`
2. Load the data using `read.csv`, and  
output the `head` of the table with `renderTable`
</div>
</div>
<div class = "col-md-2">
<div style="color:black; float:right"> `r icon::fa("clock")` 3:00 </div>
</div>
</div>

### &nbsp;

### Solution 

```{r echo = TRUE, eval = FALSE}
# ui
main <- fluidPage(
    tableOutput(
        outputId = "table"))
fluidPage(
    sidebarLayout(
        sidebarPanel = NULL,
        mainPanel = main))

# server
movies <- read.csv("movies.csv")
output$table <- renderTable(head(movies))
```

```{r}
div(tableOutput(outputId = "table"), style = "font-size:8pt")
```

```{r context = "server"}
output$table <- renderTable(head(movies), striped = TRUE, bordered = TRUE)
```

## Task 2: Basic plotting

<div class = "row">
<div class = "col-md-10">
<div style="color:red;font-weight:bold">
1. Replace the above `tableOutput` with a `plotOutput`
2. Render a scatter plot of variables  
`Body_Count` (x-axis) vs. `IMDB_Rating` (y-axis)
</div>
</div>
<div class = "col-md-2">
<div style="color:black; float:right"> `r icon::fa("clock")` 4:00 </div>
</div>
</div>

### &nbsp;

### Solution 

<div class = "row">
<div class = "col-md-6">
```{r echo = TRUE, eval = FALSE}
# ui
main <- fluidPage(
    plotOutput(
        outputId = "scatter"))
fluidPage(
    sidebarLayout(
        sidebarPanel = NULL,
        mainPanel = main))

# server
library(ggplot2)
output$scatter <- renderPlot({
    ggplot(movies, aes(
        x = Body_Count, 
        y = IMDB_Rating)) +
    geom_point()
})
```
</div>
<div class = "col-md-6">
```{r context = "server"}
output$scatter1 <- renderPlot({
    ggplot(movies, aes(x = Body_Count, y = IMDB_Rating)) +
    geom_point()
})
```

```{r}
plotOutput(outputId = "scatter1")
```
</div>
</div>

## Task 3: Reactive plotting

<div class = "row">
<div class = "col-md-10">
<div style="color:red;font-weight:bold">
1. In the `sidebarPanel`, create 2 `selectInput`s  
that enable the user to select which variables to plot
2. Modify `renderPlot` to react to the currently selected variables
</div>
</div>
<div class = "col-md-2">
<div style="color:black; float:right"> `r icon::fa("clock")` 6:00 </div>
</div>
</div>

### &nbsp;

### Solution

<div class = "row">
<div class = "col-md-6">
```{r echo = TRUE, eval = FALSE}
# ui
sidebarPanel(
    selectInput(
        intputId = "xvar", 
        label = "x-axis:", 
        choices = names(movies),
        selected = "Body_Count"),
    selectInput(
        intputId = "yvar", 
        label = "y-axis:", 
        choices = names(movies),
        selected = "IMDB_Rating"))
```
</div>
<div class = "col-md-6">
```{r echo = TRUE, eval = FALSE}
# server
output$scatter <- renderPlot({
    ggplot(movies, aes_string(
        x = input$xvar, 
        y = input$yvar)) +
    geom_point()
})
```
</div>
</div>
```{r context = "server"}
library(ggplot2)
output$scatter2 <- renderPlot({
    ggplot(movies, aes_string(x = input$xvar1, y = input$yvar1)) +
    geom_point()
})
```
<div class = "row">
<div class = "col-md-4">
```{r}
selectInput("xvar1", "x-axis:", names(movies), selected = "Body_Count")
selectInput("yvar1", "y-axis:", names(movies), selected = "IMDB_Rating")
```
</div>
<div class = "col-md-8">
```{r}
plotOutput(outputId = "scatter2")
```
</div>
</div>

## Task 4: Subsetting

<div class = "row">
<div class = "col-md-10">
<div style="color:red;font-weight:bold">
1. Add a `sliderInput` with range [0, 1000]
2. Modify `renderPlot` to plot only the subset of data  
for which the `Body_Count` lies in the specified range
3. Add a `textOutput` that prints the number of points  
that fall into the current range
</div>
</div>
<div class = "col-md-2">
<div style="color:black; float:right"> `r icon::fa("clock")` 5:00 </div>
</div>
</div>

### &nbsp;

### Solution

<div class = "row">
<div class = "col-md-6">
```{r echo = TRUE, eval = FALSE}
# ui
main <- mainPanel(
    fluidPage(
        sliderInput(
            inputId = "slider",
            label = "Range to plot:"
            min = 0, max = 1000,
            value = c(0, 100)),
    plotOutput(outputId = "scatter"),
    textOutput(outputId = "text")))

# server
library(dplyr)
output$scatter <- renderPlot({
    range <- input$slider
    subset <- filter(movies,
        Body_Count >= range[1],
        Body_Count <= range[2])
    ggplot(subset, aes_string(
        x = input$xvar, 
        y = input$yvar)) +
        geom_point()
})
output$text <- renderText({
    range <- input$slider
    subset <- filter(movies,
        Body_Count >= range[1],
        Body_Count <= range[2])
    paste("# movies:", nrow(subset))
})
```
</div>
<div class = "col-md-6">
```{r context = "server"}
output$scatter3 <- renderPlot({
    range <- input$slider1
    subset <- filter(movies, 
        Body_Count >= range[1],
        Body_Count <= range[2])
    ggplot(subset, aes_string(
        x = input$xvar2, 
        y = input$yvar2)) +
        geom_point()
})
output$text <- renderText({
    range <- input$slider1
    subset <- filter(movies, 
        Body_Count >= range[1],
        Body_Count <= range[2])
    paste("# movies:", nrow(subset))
})
```

```{r}
selectInput("xvar2", "x-axis:", names(movies), selected = "Body_Count")
selectInput("yvar2", "y-axis:", names(movies), selected = "IMDB_Rating")
sliderInput(
    inputId = "slider1", 
    label = "Body_Count range to plot:", 
    min = 0, max = 1000, value = c(0, 100))
```

```{r}
plotOutput(outputId = "scatter3")
textOutput(outputId = "text")
```
</div>
</div>

## Task 5: Using `reactive`

<div class = "row">
<div class = "col-md-10">
<div style="color:red;font-weight:bold">
1. Create reactive value to store the current subset  
of the data using `subset <- reactive()`
2. Modify `renderPlot` and `renderText` to use this reactive value
</div>
</div>
<div class = "col-md-2">
<div style="color:black; float:right"> `r icon::fa("clock")` 5:00 </div>
</div>
</div>

### &nbsp;

### Solution

<div class = "row">
<div class = "col-md-6">
```{r echo = TRUE, eval = FALSE}
# server
subset <- reactive({
    range <- input$slider
    filter(movies,
        Body_Count >= range[1],
        Body_Count <= range[2])
})
output$scatter <- renderPlot({
    ggplot(subset(), aes_string(
        x = input$xvar, 
        y = input$yvar)) +
    geom_point()
})
output$text <- renderText(
  paste("# movies:", nrow(subset())))
```
</div>
<div class = "col-md-6">
```{r context = "server"}
subset <- reactive({
    range <- input$slider2
    filter(movies,
        Body_Count >= range[1],
        Body_Count <= range[2])
})
output$scatter4 <- renderPlot(
    ggplot(subset(), aes_string(
        x = input$xvar3,
        y = input$yvar3)) +
        geom_point())
output$text2 <- renderText(paste("# movies:", nrow(subset())))
```

```{r}
selectInput("xvar3", "x-axis:", names(movies), selected = "Body_Count")
selectInput("yvar3", "y-axis:", names(movies), selected = "IMDB_Rating")
sliderInput(
    inputId = "slider2", 
    label = "Body_Count range to plot:", 
    min = 0, max = 1000, value = c(0, 100))
plotOutput(outputId = "scatter4")
textOutput(outputId = "text2")
```
</div>
</div>