---
title: "Movies app part 2"
subtitle: "SIB - Introduction to R Shiny"
author: "Phil Cheng, Helena L. Crowell"
date: "May 6th, 2019"
output: 
  learnr::tutorial:
    progress: true
    theme: flatly
runtime: shiny_prerendered
---

```{r echo = FALSE, message = FALSE, warning = FALSE}
library(dplyr)
library(ggplot2)
library(learnr)
#learnr::initialize_tutorial()
knitr::opts_chunk$set(echo = FALSE)
```

## More sophisticated apps

<div style="font-weight:bold">
The goal of this task is to

1. use an `actionButton` to add more control to our app, and
2. implement a `tabsetPanel` to combine the visualizations from session 1
3. add additional visualizations
</div>


```{r context = "data"}
movies <- read.csv(file.path("www", "movies.csv"))
```

## Task 1: Action Button 

<div class = "row">
<div class = "col-md-10">
<div style="color:red;font-weight:bold">
1. add an `actionButton` to the `sidebarPanel`  
of your Shiny app from task 3
2. update the `server` function such that  
the plot refreshes only(!) upon button-click  
(regardless of changes in the inputs)
</div>
</div>
<div class = "col-md-2">
<div style="color:black; float:right"> `r icon::fa("clock")` 3:00 </div>
</div>
</div>

### &nbsp;

### Solution

<div class = "row">
<div class = "col-md-6">
```{r echo = TRUE, eval = FALSE}
ui <- fluidPage(
  sidebarLayout(
    sidebarPanel(
      ...
      actionButton(
        inputId = "go", 
        label = "Refresh")),
    mainPanel(
      ...
    )))
```
</div>
<div class = "col-md-6">
```{r echo = TRUE, eval = FALSE}
# get x- & y-var. on button-click
xvar <- eventReactive(
  eventExpr = input$go, 
  valueExpr = input$xvar)

yvar <- eventReactive(
  eventExpr = input$go, 
  valueExpr = input$yvar)

output$scatter <- renderPlot({
    ggplot(movies, aes_string(
      x = xvar(), y = yvar())) + 
    geom_point()
})
```
</div>
</div>

<div class = "row">
<div class = "col-md-6">
```{r context = "server"}
xvar1 <- eventReactive(input$button1, input$xvar1)
yvar1 <- eventReactive(input$button1, input$yvar1)
output$scatter1 <- renderPlot(
  ggplot(movies, aes_string(x = xvar1(), y = yvar1())) + geom_point())
```
```{r}
selectInput("xvar1", "x-axis:", names(movies), selected = "Body_Count")
selectInput("yvar1", "y-axis:", names(movies), selected = "IMDB_Rating")
actionButton("button1", "Refresh")
```
</div>
<div class = "col-md-6">
```{r}
plotOutput(outputId = "scatter1")
```
</div>
</div>

## Task 2: Tabsets

<div class = "row">
<div class = "col-md-10">
<div style="color:red;font-weight:bold">
- add a `tabsetPanel` inside of the `mainPanel`:
  - Tab 1: scatter plot from task 1
  - Tab 2: `dataTableOutput` displaying the head of the table  
(task 1 in previous session)
  - Tab 3: `sliderInput` & filtered scatter plot  
(task 5 in previous session)
- inputs of both scatter plots should be shared
- 1st scatter plot should be controlled by the `actionButton`  
- 2nd scatter should be "live" / reactive to inputs
</div>
</div>
<div class = "col-md-2">
<div style="color:black; float:right"> `r icon::fa("clock")` 8:00 </div>
</div>
</div>
**Hint: Scatter `plotOutput`s should have unique IDs**

### &nbsp;

### Solution

```{r, echo = TRUE, eval = FALSE}
shinyApp(
    ui = fluidPage(
      sidebarLayout(
        sidebarPanel(...),
        mainPanel(
          tabsetPanel(
            tabPanel(
              title = "Scatter plot", 
              plotOutput(outputId = "scatter1")),
            tabPanel(
              title = "Table", 
              dataTableOutput(outputId = "table")),
            tabPanel(
              title = "Subset plot",
              sliderInput(
                intputId = "slider", 
                label = "Range to plot:", 
                min = 0, max = 1000, value = c(0, 100)),
              plotOutput(outputId = "scatter2")))
        )
      )
    ),
    server = function(input, output) {
        # load packages & data
        library(dplyr)
        library(ggplot2)
        movies <- read.csv("movies.csv")
        
        # 1st tab: `actionButton` controlled scatter plot
        xvar <- eventReactive(input$go, input$xvar)
        yvar <- eventReactive(input$go, input$yvar)
        output$scatter1 <- renderPlot(
            ggplot(movies, aes_string(xvar(), yvar())) + 
            geom_point())
        
        # 2nd tab: `dataTableOutput` displaying dataset head
        output$table <- renderDataTable(head(movies))
        
        # 3rd tab: reactive filtered scatter plot
        subset <- reactive({
            range <- input$slider
            filter(movies,
                Body_Count >= range[1],
                Body_Count <= range[2])
        })
        output$scatter2 <- renderPlot(
            ggplot(subset(), aes_string(
                x = input$xvar, 
                y = input$yvar)) +
                geom_point())
    }
)
```

## Task 3: More reactivity
<div class = "row">
<div class = "col-md-10">

<div style="font-weight:bold">
Our app currently has 4 inputs:

- 2 `selectInput`s to select columns from `movies`
- 1 `actionButton` to control tab 1's scatter plot
- 1 `sliderInput` to subset the data for tab 3's scatter plot
</div>

***

<div style="color:red;font-weight:bold">
1. Create a 4th tab that outputs:  
2 `textOutput`s and 1 `tableOutput`
2. The two `textOutput`s should show  
the mean of the respective `input$*` (unfiltered) column   
(if the column is not numeric, output `"Not a numeric"`)
3. The `tableOutput` should display the 2 selected columns  
with the rows subsetted by the `sliderInput`
4. All outputs from tab 4 should be controlled  
by the `actionButton` from task 1
5. To make navigation easier, you can  
move the `sliderInput` to the `sidebarPanel`
</div>
</div>
<div class = "col-md-2">
<div style="color:black; float:right"> `r icon::fa("clock")` 10:00 </div>
</div>
</div>

### &nbsp;

### Solution 

```{r, echo = TRUE, eval = FALSE}
shinyApp(
  ui = fluidPage(
    sidebarLayout(
      sidebarPanel(...),
      mainPanel(
        ...
        tabPanel(
          title = "Table", 
          tableOutput(outputId = "table1")), # use unique ID!
        ...
        tabPanel(
          title = "Subset Info",
          textOutput(outputId = "mean_xvar"),
          textOutput(outputId = "mean_yvar"),
          tableOutput(outputId = "table2")))
      )
    )
  ),
  server = function(input, output) {
    # load packages & data
    ...
    
    # 1st tab: `actionButton` controlled scatter plot
    ...
    
    # 2nd tab: `dataTableOutput` displaying dataset head
    ...
    
    # 3rd tab: reactive filtered scatter plot
    subset <- reactive({
        range <- input$slider
        filter(movies,
            Body_Count >= range[1],
            Body_Count <= range[2])
    })
    ...
    
    # 4th tab: `actionButton` controlled means & table subset
    
    # means of selected variables
    mx <- eventReactive(input$go, {
      u <- movies[, input$xvar]
      ifelse(is.numeric(u), mean(u), "Not a numeric")
    })
    my <- eventReactive(input$go, {
      u <- movies[, input$yvar]
      ifelse(is.numeric(u), mean(u), "Not a numeric")
    })
    output$mean_xvar <- renderText(mx())
    output$mean_yvar <- renderText(my())
    
    # filtered 2-column table
    table <- eventReactive(input$go, 
      select(subset(), input$xvar, input$yvar))
    output$table2 <- renderTable(table())
}
```

