---
title: "Deployment, debugging, resources"
subtitle: "SIB - Introduction to R Shiny"
author: "Phil Cheng, Helena L. Crowell"
date: "May 6th, 2019"
output: 
    learnr::tutorial:
    progressive: true
theme: flatly
runtime: shiny_prerendered
---
 
```{r echo = FALSE}
learnr::initialize_tutorial()
```

## Debugging

### Reactivity log

- interactive browser to visualize  
reactive dependenvies & execution

- can be displayed during...

```{r eval = FALSE}
shinyApp(ui, server)
# ctrl/command + F3 (win/mac)
```

- ...or prior to app launch

```{r eval = FALSE}
reactlogShow()
shinyApp(ui, server)
```

## Deployment/sharing

2 basic options:

1. **R script(s)**:
    - simplest way
    - user has to have R (and know how to use it)

2. **web page**:
    - most user friendly (navigation to app via web brower)
    - need a server

### 1. Sharing as R script(s)

- as R script(s) (what we've seen so far)
```{r eval = FALSE}
# single file:
runApp("<path_to_app.R>")

# multiple files:
runApp("<path_to_server.R>")
```

- as a stand-alone package
```{r eval = FALSE}
library(package)
launch_app()
```

### 2. Sharing as a web page

- via hosting a [Shiny server](https://github.com/rstudio/shiny-server/blob/master/README.md)
- via RStudio's [shinyapps.io](http://www.shinyapps.io/)  
(free with limited # of apps & active hours; can pay for more)
- ...

## Final task

- visit https://www.shinyapps.io
- click on **Sign Up** to create an account
- follow the **GETTING STARTED** instructions
- complete the skeleton script at `<script_name>`
- upload your app to your shinyapps.io account
- share the link to your app with us on the Etherpad

### Data description

- **affairs.csv** contains data from 601 men and women from the US  
who were asked how many (if any) affairs they've had in their marriage,  
along with additional questions about their personna
- The dataset contains the following variables:

variable | description        | type
:--------|:-------------------|:-----------
`sex`    | male or female     | categorical
`age`    | 17.5-57            | continuous
`ym`     | # of years married | continuous
`child`  | yes or no          | categorical
`religious`  | 1 (not) - 5 (very)  | categorical
`education`  | 9 (low) - 20 (high) | categorical
`occupation` | 1 - 7 (hollingshead classification) | categorical
`rate` | marriage self-rating; 1 (very unhappy) - 5 (very happy) | categorical
`nbaffairs` | # of affairs in the past year | continuous

### What needs to be in your Shiny app

- Create a Shiny app which has a sidebar that allows for 4 inputs
- The Shiny app should have 2 tabs in the mainPanel
- The first input should only select for catergorical columns (sex, child, religious, education, occupation, rate)
- The second input should only select for continuous columns (age, ym, nbaffairs)
- The third input will be a updateselectizeInput for the levels in the category from input1
- The fourth input will be a updateSliderInput for input2.
- Use observeEvent to allow for the input 3 and 4 update depending on your choices of input 1 and 2
- The first tab should have a boxplot of input1 vs input2
- The second tab should have a radioInput for "t-test" and "ANOVA"  and will output the text from `t.test`or `anova`